---
title: "README"
author: "Max Westphal (max.westphal@mevis.fraunhofer.de)"
output: md_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

# cases: Stratified Evaluation of Subgroup Classification Accuracy

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
`r badger::badge_devel(color="blue")`
`r badger::badge_cran_release(color="orange")`
`r badger::badge_github_actions(action="R-CMD-check")`
`r badger::badge_codecov()`
[![](https://img.shields.io/badge/preprint-arXiv-gold.svg)](https://arxiv.org/abs/2105.13469)
[![R-CMD-check](https://github.com/maxwestphal/cases/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/maxwestphal/cases/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

**cases** is an R package to simultaneously assess classification accuracy of multiple classifiers 
in several subgroups (strata). For instance, it allows to asses the accuracy of multiple candidate (index) diagnostic tests which is often measured with 

- sensitivity (accuracy in the diseased subgroup) and
- specificity (accuracy in the healthy subgroup).

A widespread goal in diagnostic accuracy studies a so-called **co-primary** analysis of these two endpoints,
i.e. to show a significant benefit (compared to some benchmark) in sensitivity **and** specificity
for **at least one** of the candidate classifiers. The package implements different methods 
for multiplicity adjustment for that purpose (e.g. Bonferroni, maxT, pairs bootstrap).


---

## Installation

You can install the development version of cases from [GitHub](https://github.com/) with:

```{r install, eval=FALSE, echo=TRUE}
# install.packages("remotes")
remotes::install_github('maxwestphal/cases', build_vignettes = TRUE)
```


---

## Usage

A vignette which explains the basic functionality of the cases package can be displayed as 
follows:
 
```{r vignette_1, eval=FALSE, echo=TRUE}
vignette(topic="package_overview", package = "cases")
```

The following vignette shows an exemplary usage of the package in the context of biomarker assessment 
and prediction model evaluation:

```{r vignette_2, eval=FALSE, echo=TRUE}
vignette(topic="example_wdbc", package = "cases")
```


---

## References

- [Westphal, Max, and Antonia Zapf. "Statistical Inference for Diagnostic Test Accuracy Studies with Multiple Comparisons." arXiv preprint arXiv:2105.13469 (2021).](https://arxiv.org/abs/2105.13469)
