---
title: "Real-world example: biomarker assessment and prediction model evaluation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Real-world example: biomarker assessment and prediction model evaluation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The goal of is this vignette is to illustrate the cases functionality by means of a real-world 
example focused on biomarker assessment and prediction model evaluation.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```


## Preparation

Load the package:

```{r setup}
## load packages:
library(readr)
library(dplyr)
library(splitstackshape)
library(glmnet)
library(cases)
```


## Introduction

We utilize the breast cancer wisconsin (diagnostic) data set.

```{r}
#?data_wdbc
data <- data_wdbc
```


```{r}
dim(data)

table(data$diagnosis) 

## Missing values?
colSums(is.na(data)) # -> no missing values
```

```{r}
summary(data)
```


```{r}
## define minimal acceptable criteria for specificity, sensitivity:
sp0 <- 0.7
se0 <- 0.9
benchmark <- c(sp0, se0)
```

## Scenario A: Biomarker assessment

```{r}
pr <- seq(0,1, 0.1) 

quantile(data$area_peak, pr) # 500, 600, 700, 800, 900 ---> area
quantile(data$compactness_peak, pr) # 0.10, 0.15, 0.20, 0.25, 0.30 ---> compactness (perimeter^2 / area - 1.0)
quantile(data$concavity_peak, pr) # 0.10, 0.15, 0.20, 0.25, 0.30 ---> concavity (severity of concave portions of the contour)

```

```{r}
cc <- c(500, 600, 700, 800, 900, 
        0.10, 0.15, 0.20, 0.25, 0.30,
        0.10, 0.15, 0.20, 0.25, 0.30) 
```


```{r}
comp_bm <- data %>% 
  dplyr::select(area_peak, compactness_peak, concavity_peak) %>% 
  cases::categorize(cc, rep(1:3, each=5)) %>% 
  cases::compare(labels = as.numeric(as.character(data$diagnosis)))

results_bm <- cases::evaluate(comp_bm, benchmark = benchmark,
                              alternative = "greater", alpha = 0.025,
                              adj="boot", regu=1) 
results_bm
```


```{r}
visualize(results_bm)
```



## Scenario B: Prediction model evaluation

```{r}
## data splitting:
set.seed(1337)
split <- stratified(data, c('diagnosis'), 1/3, bothSets = TRUE)
val <- split[[1]] %>% as.data.frame()
trn <- split[[2]] %>% as.data.frame()
dim(trn)
dim(val)
table(val$diagnosis)
```

```{r}
## train example model
mod <- glmnet(x=trn[,-1], y=trn[,1], family="binomial", alpha=0.25)
str(mod, 0)
```

```{r}
set.seed(1337)

## define hyperparameter values for L1/L2 penalty mixing parameter (alpha):
aa <- c(0, 0.25, 0.5, 0.75, 1)

## train models and create predictions:
pred_pm <- sapply(aa, function(alpha){
  mod_pm <- cv.glmnet(x = as.matrix(trn[,-1]), y=trn[,1],
                      family = "binomial",
                      type.measure = "class",
                      alpha = alpha)
  message(paste0("cv.glmnet (alpha = ", alpha, "):"))
  print(mod_pm)
  message("+++++")
  predict(mod_pm, as.matrix(val[,-1]), type="response")
})
colnames(pred_pm) <- paste0("en", aa*100)


```

```{r}
head(pred_pm)
```


```{r}
## define cutpoints (probability scale):
cc <- rep(seq(0.1, 0.5, 0.1), 5)
mm <- rep(1:5, each=5)

## create predictions: 
comp_pm <- pred_pm %>% 
  cases::categorize(cc, mm) %>% 
  cases::compare(labels = as.numeric(as.character(val$diagnosis)))
str(comp_pm, 1)
```

```{r}
## conduct statistical analysis:
set.seed(1337)
results_pm <- cases::evaluate(comp_pm, benchmark = benchmark,
                              alternative = "greater", alpha = 0.025,
                              adj="boot", regu=1) 
str(results_pm, 1)
```



```{r}
results_pm %>% lapply(filter, reject_all) 
```


```{r}
visualize(results_pm)
```

---

## References

1. Westphal M, Zapf A. Statistical inference for diagnostic test accuracy studies with multiple comparisons. Statistical Methods in Medical Research. 2024;0(0). [doi:10.1177/09622802241236933](https://journals.sagepub.com/doi/full/10.1177/09622802241236933)
- "Breast Cancer Wisconsin (Diagnostic) Data Set" at the UCI Machine Learning Repository: https://archive.ics.uci.edu/ml/datasets/breast+cancer+wisconsin+(diagnostic) 